Discovery of s-process stars in the LAMOST survey

### 2018-01-26 article/ms.tex
final draft

\section{Methods} \label{sec:methods}
\subsection{LAMOST analysis}
\subsubsection{Data-driven analysis}
The LAMOST (Large sky Area Multi-Object Fibre Spectrographic) survey released low-resolution ($\mathcal{R} \approx 1800$) optical spectra (3700\,\AA\ to 9000\,\AA) for over two million stars in their second data release \citep{luo2015}. Stellar parameters ($T_{\rm eff}$, $\log{g}$, [M/H], [$\alpha$/M]) were derived using a data-driven model constructed using high fidelity APOGEE labels of 9,952 giant stars in common between LAMOST and the APOGEE survey. It is important to note, this sample is exclusive to giant stars presenting the possibility that this research missed dwarf s-process enhanced stars. Cross-validation experience shows that the typical uncertainties are 70K in effective temperature $T_{eff}$, 0.1 dex in surface gravity log g, 0.1 dex in metallicity [Fe/H], and 0.04 dex in the alpha abundance relative to iron [$\alpha$/M]. These uncertainties are approximately comparable to the conservative APOGEE DR12 uncertainties.

@@

\begin{table*}
\centering
\caption{Available online in its entirety. Here we show a portion to demonstrate its style and content}
\label{table:table1}
\begin{tabular}{@{}|l|l|l|l|l|l|l|l|l|l|@{}}
\toprule
2MASSID             & RA (hms         & DEC (hms)        & Vr (km/s) & S/N & Teff (K) & Log g & {[}Fe/H{]} & {[}$\alpha$/H{]} & $\chi^2$  \\ \midrule
J000134.95+490743.2 & 00:01:34.96 & +49:07:43.2 & -42.27    & 72  & 5044 & 3.11  & -0.54      & 0.11      & 0.79 \\ \midrule
J000403.80+160257.1 & 00:04:03.80 & +16:02:57.2 & -35.38    & 42  & 5200 & 3.40  & -0.41      & 0.09      & 0.33 \\ \midrule
J000908.48+421615.4 & 00:09:08.49 & +42:16:15.5 & -34.78    & 36  & 4825 & 3.02  & -0.46      & 0.16      & 0.25 \\ \midrule
J001754.17+394140.0 & 00:17:54.17 & +39:41:40.0 & -42.87    & 39  & 5008 & 3.16  & -0.64      & 0.12      & 0.23 \\ \midrule
J001841.57+040533.7 & 00:18:41.57 & +04:05:33.8 & -20.69    & 141 & 4840 & 3.68  & -0.52      & 0.06      & 0.93 \\ \midrule
J002107.60+361533.0 & 00:21:07.60 & +36:15:33.1 & -53.06    & 44  & 4854 & 2.87  & 0.01       & 0.06      & 0.54 \\ \midrule
J002537.59+411123.7 & 00:25:37.59 & +41:11:23.8 & -12.59    & 36  & 5075 & 3.30  & -0.32      & 0.10      & 0.33 \\ \midrule
J003755.55+571706.8 & 00:37:55.55 & +57:17:06.9 & -5.70     & 49  & 4046 & 1.14  & -0.09      & -0.01     & 0.83 \\ \midrule
J003813.23+423025.9 & 00:38:13.23 & +42:30:25.9 & -40.17    & 36  & 4475 & 2.12  & 0.14       & 0.00      & 0.51 \\ \midrule
J003959.14+394614.6 & 00:39:59.14 & +39:46:14.6 & 5.10      & 35  & 4955 & 3.23  & -0.40      & 0.11      & 0.28 \\ \midrule
J004119.68+324627.8 & 00:41:19.68 & +32:46:27.9 & 4.50      & 122 & 4861 & 2.59  & -0.31      & 0.08      & 1.67 \\ \midrule
J004156.56+390945.6 & 00:41:56.57 & +39:09:45.6 & -254.52   & 44  & 4771 & 2.34  & -0.74      & 0.16      & 0.57 \\ \midrule
J004743.01+433358.1 & 00:47:43.02 & +43:33:58.2 & -62.66    & 65  & 4834 & 2.64  & -0.60      & 0.17      & 1.06 \\ \midrule
J005101.98+314045.6 & 00:51:01.98 & +31:40:45.7 & -23.68    & 98  & 5202 & 3.32  & -0.53      & 0.10      & 0.67 \\ \midrule
J005254.69+203554.7 & 00:52:54.69 & +20:35:54.7 & 35.68     & 113 & 5026 & 3.17  & -0.82      & 0.20      & 1.01 \\ \midrule
J010836.26+444435.9 & 01:08:36.26 & +44:44:35.9 & 12.59     & 49  & 5189 & 3.19  & -0.56      & 0.10      & 0.98 \\ \midrule
J011919.20+195227.0 & 01:19:19.21 & +19:52:27.0 & -35.38    & 31  & 4897 & 2.77  & -0.54      & 0.11      & 0.29 \\ \midrule
J011930.90-015027.3 & 01:19:30.91 & -01:50:27.4 & -143.30   & 44  & 5040 & 3.22  & -0.91      & 0.20      & 0.48 \\ \midrule
J020316.40+462211.1 & 02:03:16.41 & +46:22:11.1 & 11.09     & 37  & 4897 & 2.89  & -0.39      & 0.11      & 0.51 \\ \bottomrule
\end{tabular}
\end{table*}

\subsubsection{Enhancements due to sodium, technetium, and carbon}
We performed an identical analysis to the process described in Section \ref{sec:cand} in order to identify sodium, technetium, or carbon enhancement in our sample of 257 Barium stars. We used the sodium doublet lines at 5889\,\AA\ and 5895\,\AA\ to select Barium stars enhanced in sodium, and required significant absorption in both features. Only 5/257 Barium stars met this criteria,  \todo{however something about the dust for those stars}. For technetium enhancement we searched for significant residual deviations at the 4049\,\AA, 4238\,\AA, 4262\,\AA, 4297\,\AA, and 5924\,\AA. These absorption lines are extremely weak, and would require a substantial amount of technetium before it would be visible in a high S/N LAMOST spectrum. We found 51 Barium stars exhibited some level of significant enhancement (based on the filters above) at a single technetium line, yet none showed enhancement at more than one technetium line. We note that the single enhancement near technetium for these 51 stars may be the consequence of data reduction/calibration artefacts, but we include details of which stars matched on technetium in Table \ref{tab:table} nonetheless.
Finally, for carbon enhancement we searched for significant deviations at the CH and G band near 4300\,\AA. We found 178/257 stars to show significant carbon enhancement (e.g., $[\textrm{C/M}] \gtrsim 0.5$).

@@

We find only $<2$\,\% of our Barium stars show sodium enhancement, contrary to previous studies \citep[e.g.][]{decastro2016}. We also note that the 5 sodium-enhanced Barium stars all have $\log{g} \approx 3$, not at lower $\log{g}$ values as might be expected if low $\log{g}$ values systematically contributed to higher [Na/Fe] abundance ratios. These results suggest that sodium-rich material from a companion TP-AGB star has polluted these 5 Barium candidates, but importantly, sodium enhancement is by no means ubiquitous in Barium stars. It's important to note, sodium enhancement can occur due to the presence of interstellar dust, and using the IRSA's all-sky dust map \citep{schlafly2011}, E(B-V) values were calculated for our 5 candidates. The results show 2 sodium enhanced stars have strong ($\approx$0.35) E(B-V) values, higher then galactic stellar levels, suggesting these candidates are false positives. Whereas, 3 exhibit low ($\approx$0.045) E(B-V) values, validating our sodium enhancement results for these candidates. The comparisons made in \citep{decastro2016} made use of multiple literature comparisons, and therefore it is possible that there are systematic differences (biases) between Milky Way studies of normal FGK-type stars and those focussed on Barium stars, leading to unresolved differences in the reported [Na/Fe] abundance ratio. These same systematic effects could contribute to the observed correlation between [Na/Fe] and $\log{g}$, which we do not observe in our sample.

>>>>

\section{Methods} \label{sec:methods}
\subsection{LAMOST analysis}
\subsubsection{Data-driven analysis}
The LAMOST (Large sky Area Multi-Object Fibre Spectrographic) survey released low-resolution ($\mathcal{R} \approx 1800$) optical spectra (3700\,\AA\ to 9000\,\AA) for over two million stars in their second data release \citep{luo2015}. Stellar parameters ($T_{\rm eff}$, $\log{g}$, [M/H], [$\alpha$/M]) were derived using a data-driven model constructed using high fidelity APOGEE labels of 9,952 giant stars in common between LAMOST and the APOGEE survey. It is important to note, this sample is exclusive to giant stars presenting the possibility that this research missed dwarf s-process enhanced stars. Cross-validation experience shows that the typical uncertainties are 70K in effective temperature $\rm T_{\rm eff}$, 0.1 dex in surface gravity log g, 0.1 dex in metallicity [Fe/H], and 0.04 dex in the alpha abundance relative to iron [$\alpha$/M]. These uncertainties are approximately comparable to the conservative APOGEE DR12 uncertainties.

@@

\begin{table*}
\centering
\caption{Available online in its entirety. Here we show a portion to demonstrate its style and content}
\label{table:table1}
\begin{tabular}{@{}|l|l|l|l|l|l|l|l|l|l|@{}}
\toprule
2MASSID             & RA (hms         & DEC (hms)        & Vr (km/s) & S/N & $\rm T_{\rm eff}$ (K) & Log g & {[}Fe/H{]} & {[}$\alpha$/H{]} & $\chi^2$  \\ \midrule
J000134.95+490743.2 & 00:01:34.96 & +49:07:43.2 & -42.27    & 72  & 5044 & 3.11  & -0.54      & 0.11      & 0.79 \\ \midrule
J000403.80+160257.1 & 00:04:03.80 & +16:02:57.2 & -35.38    & 42  & 5200 & 3.40  & -0.41      & 0.09      & 0.33 \\ \midrule
J000908.48+421615.4 & 00:09:08.49 & +42:16:15.5 & -34.78    & 36  & 4825 & 3.02  & -0.46      & 0.16      & 0.25 \\ \midrule
J001754.17+394140.0 & 00:17:54.17 & +39:41:40.0 & -42.87    & 39  & 5008 & 3.16  & -0.64      & 0.12      & 0.23 \\ \midrule
J001841.57+040533.7 & 00:18:41.57 & +04:05:33.8 & -20.69    & 141 & 4840 & 3.68  & -0.52      & 0.06      & 0.93 \\ \midrule
J002107.60+361533.0 & 00:21:07.60 & +36:15:33.1 & -53.06    & 44  & 4854 & 2.87  & 0.01       & 0.06      & 0.54 \\ \midrule
J002537.59+411123.7 & 00:25:37.59 & +41:11:23.8 & -12.59    & 36  & 5075 & 3.30  & -0.32      & 0.10      & 0.33 \\ \midrule
J003755.55+571706.8 & 00:37:55.55 & +57:17:06.9 & -5.70     & 49  & 4046 & 1.14  & -0.09      & -0.01     & 0.83 \\ \midrule
J003813.23+423025.9 & 00:38:13.23 & +42:30:25.9 & -40.17    & 36  & 4475 & 2.12  & 0.14       & 0.00      & 0.51 \\ \midrule
J003959.14+394614.6 & 00:39:59.14 & +39:46:14.6 & 5.10      & 35  & 4955 & 3.23  & -0.40      & 0.11      & 0.28 \\ \midrule
J004119.68+324627.8 & 00:41:19.68 & +32:46:27.9 & 4.50      & 122 & 4861 & 2.59  & -0.31      & 0.08      & 1.67 \\ \midrule
J004156.56+390945.6 & 00:41:56.57 & +39:09:45.6 & -254.52   & 44  & 4771 & 2.34  & -0.74      & 0.16      & 0.57 \\ \midrule
J004743.01+433358.1 & 00:47:43.02 & +43:33:58.2 & -62.66    & 65  & 4834 & 2.64  & -0.60      & 0.17      & 1.06 \\ \midrule
J005101.98+314045.6 & 00:51:01.98 & +31:40:45.7 & -23.68    & 98  & 5202 & 3.32  & -0.53      & 0.10      & 0.67 \\ \midrule
J005254.69+203554.7 & 00:52:54.69 & +20:35:54.7 & 35.68     & 113 & 5026 & 3.17  & -0.82      & 0.20      & 1.01 \\ \midrule
J010836.26+444435.9 & 01:08:36.26 & +44:44:35.9 & 12.59     & 49  & 5189 & 3.19  & -0.56      & 0.10      & 0.98 \\ \midrule
J011919.20+195227.0 & 01:19:19.21 & +19:52:27.0 & -35.38    & 31  & 4897 & 2.77  & -0.54      & 0.11      & 0.29 \\ \midrule
J011930.90-015027.3 & 01:19:30.91 & -01:50:27.4 & -143.30   & 44  & 5040 & 3.22  & -0.91      & 0.20      & 0.48 \\ \midrule
J020316.40+462211.1 & 02:03:16.41 & +46:22:11.1 & 11.09     & 37  & 4897 & 2.89  & -0.39      & 0.11      & 0.51 \\ \bottomrule
\end{tabular}
\end{table*}

\subsubsection{Enhancements due to sodium, technetium, and carbon}
We performed an identical analysis to the process described in Section \ref{sec:cand} in order to identify sodium, technetium, or carbon enhancement in our sample of 257 Barium stars. We used the sodium doublet lines at 5889\,\AA\ and 5895\,\AA\ to select Barium stars enhanced in sodium, and required significant absorption in both features. Only 5/257 Barium stars met this criteria, and it's important to note sodium enhancement can occur due to the presence of interstellar dust. Using the IRSA's all-sky dust map \citep{schlafly2011}, E(B-V) values were calculated for our 5 candidates. The results show 2 sodium enhanced stars have strong ($\approx$0.35) E(B-V) values, higher then galactic stellar levels, suggesting these candidates are false positives. Whereas, 3 exhibit low ($\approx$0.045) E(B-V) values, confirming sodium enhancement results for these candidates. For technetium enhancement we searched for significant residual deviations at the 4049\,\AA, 4238\,\AA, 4262\,\AA, 4297\,\AA, and 5924\,\AA. These absorption lines are extremely weak, and would require a substantial amount of technetium before it would be visible in a high S/N LAMOST spectrum. We found 51 Barium stars exhibited some level of significant enhancement (based on the filters above) at a single technetium line, yet none showed enhancement at more than one technetium line. We note that the single enhancement near technetium for these 51 stars may be the consequence of data reduction/calibration artefacts, but we include details of which stars matched on technetium in Table \ref{tab:table} nonetheless.
Finally, for carbon enhancement we searched for significant deviations at the CH and G band near 4300\,\AA. We found 178/257 stars to show significant carbon enhancement (e.g., $[\textrm{C/M}] \gtrsim 0.5$).

@@

We find only $<2$\,\% of our Barium stars show sodium enhancement, contrary to previous studies \citep[e.g.][]{decastro2016}. We also note that the 5 sodium-enhanced Barium stars all have $\log{g} \approx 3$, not at lower $\log{g}$ values as might be expected if low $\log{g}$ values systematically contributed to higher [Na/Fe] abundance ratios. These results suggest that sodium-rich material from a companion TP-AGB star has polluted these 5 Barium candidates, but importantly, sodium enhancement is by no means ubiquitous in Barium stars. It's important to note, sodium enhancement can occur due to the presence of interstellar dust, and using the IRSA's all-sky dust map \citep{schlafly2011}, E(B-V) values were calculated for our 5 candidates. The results show 2 sodium enhanced stars have strong ($\approx$0.35) E(B-V) values, higher then galactic stellar levels, suggesting these candidates are false positives. Whereas, 3 exhibit low ($\approx$0.045) E(B-V) values, confirming sodium enhancement results for these candidates. The comparisons made in \citep{decastro2016} made use of multiple literature comparisons, and therefore it is possible that there are systematic differences (biases) between Milky Way studies of normal FGK-type stars and those focussed on Barium stars, leading to unresolved differences in the reported [Na/Fe] abundance ratio. These same systematic effects could contribute to the observed correlation between [Na/Fe] and $\log{g}$, which we do not observe in our sample.
